---
title: "FinalProject"
author: "Ben Tankus"
date: "3/2/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(Amelia)
library(broom)
library(faraway)
library(ggplot2)
library(dplyr)
library(pairsD3)
library(tidyverse)

```

```{r Initialize}
df <- read.csv('OR_acs_house_occ.csv')

```

# QUESTIONS TO ANSWER:

1. Do people living in apartments pay less on electricity than those living in houses? (adjust for number of bedrooms and number of occupants)
2. How much does each group pay? 
3. What about extra heating cost using fuel or gas?

# CREATE MODEL
Model is used to predict electricity costs for a house in Oregon

## MISSINGNESS

### ACR and VALP both have many null values, 2586 and 4632 respectively. Analysis in the *TBD* section will determine how the missingness is distributed.

```{r MISSINGNESS}
i = 1
names <- names(df)
  
for( col in df) {
  #print(paste(col.names(), sum(is.na(col))))
  print(paste(names[i], "Has", sum(is.na(col)), "nulls"))
  i <- i + 1
}



dfNulls <- df[, c('ACR','ELEP', 'VALP')]

dfNulls_long <- gather(dfNulls, variable, value, -ELEP)


qplot(ELEP, variable, data = dfNulls_long, geom = "tile",
fill = is.na(value)) +
scale_fill_manual("missing?",
values = c('TRUE' = "black", 'FALSE' = "white")) +
theme(axis.text.x = element_text(angle = 0))


VALPNullsHist.df <- subset(df[,c('VALP', 'ELEP')], is.na(VALP) == TRUE) 
hist(VALPNullsHist.df[,'ELEP'], main = 'VALP Missingness Histogram', xlab = 'ELEP')

VALPNoNullsHist.df <- subset(df[,c('VALP', 'ELEP')], is.na(VALP) == FALSE) 
qplot(VALPNoNullsHist.df[,'VALP'], VALPNoNullsHist.df[,'ELEP'], main = 'VALP VS ELEP Scatter', ylab = 'ELEP', xlab = 'VALP')



ACRNullsHist.df <- subset(df[,c('ACR', 'ELEP')], is.na(ACR) == TRUE) 
hist(ACRNullsHist.df[,'ELEP'], main = 'ACR Missingness Histogram', xlab = 'ELEP')
ACRNoNullsHist.df <- subset(df[,c('ACR', 'ELEP')], is.na(ACR) == FALSE) 
qplot(as.numeric(factor(ACRNoNullsHist.df[,'ACR'])), ACRNoNullsHist.df[,'ELEP'], main = 'ACR VS ELEP Scatter', ylab = 'ELEP', xlab = 'ACR')

```
```{r}
panel.hist <- function(x, ...) {
  usr <- par("usr")
  on.exit(par(usr))
  par(usr = c(usr[1:2], 0, 1.5) )
  h <- hist(x, plot = FALSE)
  breaks <- h$breaks
  nB <- length(breaks)
  y <- h$counts
  y <- y/max(y)
  rect(breaks[-nB], 0, breaks[-1], y, col = "white", ...)
}


panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...) {
  usr <- par("usr")
  on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- abs(cor(x, y, use = "complete.obs"))
  txt <- format(c(r, 0.123456789), digits = digits)[1]
  txt <- paste(prefix, txt, sep = "")
  if (missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
  text(0.5, 0.5, txt, cex =  cex.cor * (1 + r) / 2)
}

```


### Correlation check before dropping potentially practically irrelevant columns
```{r CORR CHECK}

df.corr <-df[,c('ELEP', 'NP', 'BDSP','BLD' ,'FULP', 'GASP', 'HFL', 'RMSP', 'TEN', 'YBL', 'R18', 'R60')]

df.corr['BLD'] = as.numeric(factor(df.corr[,'BLD']))
df.corr['HFL'] = as.numeric(factor(df.corr[,'HFL']))
df.corr['TEN'] = as.numeric(factor(df.corr[,'TEN']))
df.corr['YBL'] = as.numeric(factor(df.corr[,'YBL']))
df.corr['R18'] = as.numeric(factor(df.corr[,'R18']))
df.corr['R60'] = as.numeric(factor(df.corr[,'R60']))



#plot(df.corr)
summary(df.corr)


#pairs(df.corr, 
#      main = "Anderson's Iris Data -- 3 species",
#      pch = 21, 
#      bg = c("#1b9e77", "#d95f02", "#7570b3"))#[unclass(df.corr$Species)])

pdf(file = "pairsplot.pdf")
pairs(
  df.corr,
  upper.panel = panel.cor,
  diag.panel  = panel.hist,
  lower.panel = panel.smooth,
  gap = 1/5,
  pch = 20 #small dots
)
dev.off()

#shinypairs(df.corr)
#pairsD3(df.corr[,c(1,2,3,4,5)], opacity = 0.9, cex = 4, width = 600 )

#DECLARE FINAL RELEVANT DF
df.relevant <- df[,c('ELEP', 'FULP', 'GASP', 'YBL', 'BLD', 'HFL', 'RMSP', 'NP')]
head(df.relevant)
```


### CLEANING FINAL COLUMN CHOICES
``` {r FILTER FOR HEAT TYPE}

df.final <- df.relevant

# Drop rows with "Mobile home or trailer
df.final <- filter(df.final, BLD != 'Mobile home or trailer' & BLD != 'Boat, RV, van, etc.')

# Transform BLD
adjustVector <- c()
i <- 1

for (e in df.final[,'BLD'])
{
  
  if (grepl('house' ,e, fixed = TRUE)){
    adjustVector[i] <- "House"
    i <- i+1
    
  } else 
  {
    adjustVector[i] <- "Apartment"
    i <- i+1
  }
}
df.final['BLDAdjusted'] = adjustVector


# Transform HFL
adjustVector <- c()
i <- 1

for (e in df.final[,'HFL'])
{
  
  if (grepl('Electricity' ,e, fixed = TRUE)){
    adjustVector[i] <- "Electricity"
    i <- i+1
    
  } else 
  {
    adjustVector[i] <- "Not Electricity"
    i <- i+1
  }
}
df.final['HFLAdjusted'] = adjustVector


# TRANSFORM YBL
adjustVector <- c()
i <- 1

for (e in df.final[,'YBL'])
{
  
  if (e >= 2005){
    adjustVector[i] <- "2005 to 2015"
    i <- i+1
    
  } else 
  {
    adjustVector[i] <- e
    i <- i+1
  }
}
df.final['YBLAdjusted'] = adjustVector







```




## EXPLORE DATA


```{r EXPLORE DATA}

# NO SIGNIFICANT VISUAL CORRELATION BETWEEN COST AND NUMBER OF PERSONS
plot(df$NP, df$ELEP, main = 'Number of Persons by Electricity Cost', xlab =  'Persons', ylab = 'Cost') 

# NO SIGNIFICANT VISUAL CORRELATION BETWEEN COST AND GAS PRICS
plot(df$GASP, df$ELEP, main = 'Gas Monthly Cost by Electricity Cost', xlab =  'Gas Cost', ylab = 'Cost') 

# MAYBE CORRELATION
plot(df$FULP, df$ELEP, main = 'Electricity Monthly Cost by Fuel Cost', xlab =  'Fuel Cost', ylab = 'Cost') 

# MAYBE CORRELATION
plot(df$BDSP, df$ELEP, main = 'Electricity Monthly Cost by Bedroom Count', xlab =  'Bedroom Count', ylab = 'Cost') 

# MAYBE CORRELATION
plot(df$HFL, df$ELEP, main = 'Electricity Monthly Cost by Bedroom Count', xlab =  'Bedroom Count', ylab = 'Cost') 



```

```{r FIT MODELS}

# DECLARE HOUSE MODEL
df.house <- filter(df.final, BLDAdjusted == 'House')
fit.house <- lm(ELEP ~ factor(HFLAdjusted) + NP + RMSP - 1, data = df.house)
summary(fit.house)

# DECLARE APARTMENT MODEL
df.apt <- filter(df.final, BLDAdjusted == 'Apartment')
fit.apt <- lm(ELEP ~ factor(HFLAdjusted) + NP + RMSP - 1, data = df.apt)
summary(fit.apt)

apt.fitted <- summary(fit.apt$fitted.values)
house.fitted <- summary(fit.house$fitted.values)

apt.fitted
house.fitted


qplot(apt.fitted, geom = "boxplot", main = 'Apartment Boxplot', xlim = c(25,290), xlab = "Electricity Value")  +  theme(axis.text.y = element_blank())
qplot(house.fitted, geom = "boxplot", main = 'House Boxplot', xlim = c(25,290)) +  theme(axis.text.y = element_blank())



#str(fit.apt)


#anova(fit.house, fit.apt)

```



