---
title: "Module 1 Lab"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---

```{r, include = FALSE}
library(learnr)
library(ggplot2)
library(Sleuth3)
knitr::opts_chunk$set(echo = TRUE, exercise.eval = FALSE, warning = FALSE)
```

## Introduction

This lab explores some tools for manipulating data frames to get group summaries, both graphical and numeric. These tools facilitate exploratory analysis, as part of the process of comparing groups.

To learn these tools, we use the the *Diet Restriction and Longevity* study stored in `Sleuth3` as `case0501`. Run `?case0501` to read more about the study.

```{r, warning=FALSE}
head(case0501) # First few rows of data
str(case0501) # Structure of case0501 data object
```

Recall that `str()` is great way to get a compact summary of an R data object. It returns the class of the object, the number of observations and variables, the class of those variables, and the first few values of each variable. As you can see (along with reading `?case0501`), each row in the `case0501` data frame corresponds to one mouse, and contains information about which of six `Diet` treatments (`Factor w/ 6 levels`) that mouse was assigned, and its `Lifetime` is reported in months (`numeric`).

## Graphical Group Summaries

```{r, warning=FALSE}
qplot(x = Diet, y = Lifetime, data = case0501)
```

This plot helps us get an overall sense of the data we are studying. The diet treatments are on the x-axis (the first argument to `qplot()`), and the lifetimes are on the y-axis (the second argument to `qplot()`).  It's just a scatter plot where one of or variables happens to be categorical. 

With this plot we get a general sense of the location and spread of lifetimes within each diet. We can clearly see mice on some diets tends to have longer lifetimes (N/R40 for example). A drawback to this plot is the problem of overlapping points. For example, if five mice have the same lifetime in the same treatment group, they will only look like one mouse! One solution to this problem is introduce a little random noise into the observations: a process called jittering.  In `ggplot2` this is handled by changing the geom to `"jitter"`:

```{r}
qplot(Diet, Lifetime, data = case0501, 
  geom = "jitter")
```

This plot now gives us a better picture of how many mice are in the areas of high density, but of course the locations are no longer exact. 

Another alternative is the boxplot.

```{r}
qplot(Diet, Lifetime, data = case0501, geom = "boxplot")
```

Now each treatment group is represented by a boxplot. This is a helpful summary since it
directly presents a measure of center and spread, but be aware we necessarily lose some information by reducing our data to this smaller set of summary statistics.

In some cases it might be nice to see these plots ordered by increasing medians.

```{r}
qplot(reorder(Diet, Lifetime, FUN = median), 
  Lifetime, data = case0501, geom = "boxplot")
```

The `reorder()` function put the `Diet` boxplots in order of increasing median `Lifetime`. 

Sometimes a histogram is preferred to visually convey summary information. The following code constructs one histogram for each diet.

```{r, message=FALSE, fig.height=6}
qplot(Lifetime, data = case0501) + facet_wrap(~ Diet, ncol = 1)
```

Note that x-axis is fixed across all diets, which makes it easy to compare the spread and location of the lifetimes by diet type. In this code `qplot` automatically defines the bin width to range/30. Keep in mind that bin widths can drastically affect the appearance of a histogram.

These tools give an excellent visual understanding of the data, but we also want exact numeric characteristics.

## Numerical Group Summaries

The next line calculates the mean lifetime of each diet. 

```{r}
tapply(case0501$Lifetime, case0501$Diet, mean) # Group means
```

The function `tapply()` applies a function to groups of data. The function takes 3 arguments: (i) the first argument specifies the vector we want to apply the function to, in this case the `Lifetime` column in `case0501`, (ii) the second argument is the vector which provides the grouping variable (i.e. the categorical variable) in this case the `Diet` column in `case0502`, and (iii) the third argument specifies the function we want to apply,  in this case `mean()`. 

Typing `case0501$` repeatedly can get tiresome, so one way to avoid it is to wrap our call to `tapply()` in the function `with()`.  The first argument to `with()` specifies where to look for the variables in the second argument (in this case our `tapply()` call without all the `case0501$`s):
```{r}
with(case0501, tapply(Lifetime, Diet, mean)) # Group means
```

We can easily swap out `mean` for other functions to get other summaries:
```{r}
with(case0501, tapply(Lifetime, Diet, sd)) # standard deviations by group
n_by_group <- with(case0501, tapply(Lifetime, Diet, length)) # Group sizes
n_by_group
```

Notice that the group sizes are saved as `n_by_group`, which can be very useful for later operations.

Aside: You want to be a little cautious with that last example, `length()` tells you how many rows are in the corresponding group, but they could potentially contain missing values (`NA`), which we wouldn't want to neccessarily count.  A safer way might be:

```{r}
n_by_group <- with(case0501, 
  tapply(Lifetime, Diet, function(x) sum(!is.na(x)))) 
```

## One-way F-test

This section contains the code from Lecture 2 of Module 1, which performs the one-way ANOVA F-test.  First, we again review the structure of the data.
```{r}
head(case0501)
str(case0501)
```

```{r}
qplot(Diet, Lifetime, data=case0501)
```

Next we define the number of groups, `I`, and the total number of observations, `N`:

```{r}
I <- length(unique(case0501$Diet))
N <- nrow(case0501)
```

Then we perform the computations to find the overall mean (stored as `case0501$overall_mean`) and the group means (stored as `case0501$group_mean`):

```{r}
case0501$overall_mean <- with(case0501, mean(Lifetime))
case0501$group_mean <- with(case0501, ave(Lifetime, Diet))
```

Using the overall mean and the group mean, we can then compute the within-group sum-of-squares (`within_group_SS`) and the between-groups sum-of-squares (`between_group_SS`).

```{r}
within_group_SS <- with(case0501, sum((Lifetime - group_mean)^2))
between_group_SS <- with(case0501, sum((group_mean - overall_mean)^2))
```

Finally, we compute the F-statistic:

```{r}
F_stat <- (between_group_SS/(I-1))/(within_group_SS/(N - I))
F_stat
```

The p-value is then obtained using the `pf()` function, where the first argument is the value of the F-statistic we just computed, the second argument is the numerator degrees of freedom (`I-1`, which is the number of groups minus one), and the third argument is the denominator degrees of freedom (`N-1`, which is the total sample size minus one).

```{r}
1 - pf(F_stat, I-1, N-1)
```

### Have R do it!

Instead of doing all of these computations by hand, we could just have R do it for us using the `oneway.test()` function.  First, you can read the help file for this function by typing:

```{r}
help(oneway.test)
```

This function takes several arguments: the first argument is a *formula* that specifies the outcome variable first (in this case, the `Lifetime` variable), then a tilde (`~`), and then the grouping variable (in this case, the `Diet` variable).  We also have to tell the function where to look for these variables---that's what the `data = ` argument does. 

Finally, we have the choice of performing the standard one-way ANOVA, which assumes that the population variances are all equal, or we can perform the Welch-Satterthwaite one-way ANOVA, which modifies the statistic and reference F-distribution to account for possibly unequal population variances.  To perform standard one-way ANOVA (again, which assumes all population variances are equal), we use the `var.equal = T` argument.

```{r}
oneway.test(Lifetime ~ Diet, data = case0501, var.equal = T)
```

Note that the output here matches the test statistic value and the p-value that we computed by hand.

Next we can instead do the Welch-Satterthwaite one-way ANOVA, using the `var.equal = F` argument:

```{r}
# Try without assuming equal variance
oneway.test(Lifetime ~ Diet, data = case0501, var.equal = F)
```

The test statistic changes, but the p-value is still *very* small!  In other datasets you might find that the p-values are very different if you use the `var.equal=T` option or the `var.equal=F` option.

## Kruskal-Wallace

R also has a function `kruskal.test()` that performs the Kruskal-Wallace Test.  It has very similar structure and arguments to the `oneway.test()` function:

```{r}
kruskal.test(Lifetime ~ Diet, data = case0501)
```

The output of this function is also similar to the output of the `oneway.test()` function: it provides a test statistic value (here the value of the test statistic is 159.01) and a p-value (here the value of the p-value is effectively 0).

Remember that the results of the Kruskal-Wallace test *will not* always match the results of the one-way ANOVA: *these tests are answering different questions!*  