---
title: "Module 5 Lab Submission"
author: "Ben Tankus"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(broom)
library(ggplot2)
library(MASS)
library(Sleuth3)
```

Again we consider the Brain size data in the data set `case0902` from the `Sleuth3` library. You can read more about this data set by viewing the help file:

```{r echo=TRUE}
help(case0902)
head(case0902)
```

In Lecture 2 of this Module, we considered modeling all four variables on the log scale, and we compared two models: a full model with log-Body, log-Gestation, and log-Litter as predictors, and a reduced model with only log-Body as a predictor.

1. **Fit the two models described above (and as shown on slide 15 of Lecture 2).**

```{r echo=TRUE}
fitFull <- lm(log(Brain) ~ log(Body) + log(Gestation) + log(Litter), data = case0902)
fitRed <- lm(log(Brain) ~ log(Body), data = case0902)


```

Now suppose we have four new species, with the following values of Body, Gestation, and Litter:

| Species | Body | Gestation | Litter |
|:-------:|:----:|:---------:|:------:|
| A       | 1    | 50        | 1      |
| B       | 1    | 50        | 2      |
| C       | 1    | 150       | 1      |
| D       | 50   | 50        | 1      |


2. **Create a new data frame with these values, and use the `predict()` function to predict the log-brain size for these new species.**

```{r echo=TRUE}
newData = data.frame(
  Species = c('A', 'B', 'C', 'D')
  , Body = c(1,1,1,50)
  , Gestation = c(50,50,150,50)
  , Litter = c(1,2,1,1)
)
print('FULL MODEL')
predict(fitFull, newdata = newData)
print('REDUCED MODEL')
predict(fitRed, newdata = newData)


```


3. **Obtain confidence intervals for the mean log-brain size of the new species above, using the `interval = "confidence"` argument in the `predict()` function.**

```{r}
   
print('FULL MODEL')
predict(fitFull, newdata = newData, interval = 'confidence')
print('REDUCED MODEL')
predict(fitRed, newdata = newData, interval = 'confidence')



```


4. **Obtain prediction intervals for the log-brain size of the new species above, using the `interval = "prediction"` argument in the `predict()` function.**

```{r}
   
print('FULL MODEL')
predict(fitFull, newdata = newData, interval = 'prediction')
print('REDUCED MODEL')
predict(fitRed, newdata = newData, interval = 'prediction')


```


5. **Compare the reduced model to the full model using the AIC and BIC measures.**

```{r}

anova(fitFull, fitRed) #ANOVA SHOWS SIGNIFICANT DIFFERENCE

AIC(fitFull,fitRed)
BIC(fitFull,fitRed)
   
```


## In both AIC and BIC the **full model** is superior and has a lower value than the reduced model.