---
title: "Module 6 Lab Submission"
author: "Ben Tankus"
output: pdf_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Amelia)
library(broom)
library(faraway)
library(ggplot2)
library(Sleuth3)
library(tidyverse)
```


Again we consider the Brain size data in the data set `case0902` from the `Sleuth3` library. You can read more about this data set by viewing the help file:

```{r echo=TRUE}
help(case0902)
head(case0902)
```

In the previous Module, we considered a model using the log scale for  all four variables, where the goal is to model log-brain-size as a function of log-gestation, log-littersize, and log-body-size.  Here, we are going to arbitrarily set some of those predictor values to be missing:

```{r}
set.seed(123)
brainMiss <- case0902
n <- nrow(brainMiss)
brainMiss$Body[sample(1:n, 10, replace=FALSE)] <- NA
brainMiss$Litter[sample(1:n, 7, replace=FALSE)] <- NA
brainMiss$Gestation[sample(1:n, 15, replace=FALSE)] <- NA
```

1. **Create a visual representation of the missingness pattern in this data, using one of the methods demonstrated in the lab example.**

```{r}
brainMiss$zip <- rownames(brainMiss)
brainMiss_long <- gather(brainMiss, variable, value, -zip)
(brainMiss_long)


qplot(zip, variable, data = brainMiss_long, geom = "tile", 
fill = is.na(value)) +
scale_fill_manual("missing?", 
values = c(`TRUE` = "black", `FALSE` = "white")) +
theme(axis.text.x = element_text(angle = 90))

```



2. **Use the `amelia` function to perform multiple imputation for these missing values, based on `m=50` imputed datasets.**

```{r echo=TRUE}
brainMiss <- case0902
n <- nrow(brainMiss)
brainMiss$Body[sample(1:n, 10, replace=FALSE)] <- NA
brainMiss$Litter[sample(1:n, 7, replace=FALSE)] <- NA
brainMiss$Gestation[sample(1:n, 15, replace=FALSE)] <- NA


brainMissQuant <- subset(brainMiss, select = -Species)

n.imp <- 50
ame.imp <- amelia(brainMissQuant, m=n.imp, p2s = 0, log = 2:4)


```

3. **Fit models using each of the imputed datasets**

```{r echo=TRUE}
betas <- matrix(0, nrow=n.imp, ncol=4)
ses <- matrix(0, nrow=n.imp, ncol=4)

for(i in 1:n.imp){
newMod <- lm(log(Brain) ~ log(Body) + log(Gestation) + log(Litter), data=ame.imp$imputations[[i]])
betas[i,] <- coef(newMod)
14
ses[i,] <- coef(summary(newMod))[,2]
}


```

4. **Use the `mi.meld` function to find the multiple imputation estimates of the coefficients.**

```{r echo=TRUE}

mi.meld(q = betas, se = ses)

```

Recall the coefficients from the model constructed on the full data (with no missing values):

```{r}
origMod <- lm(log(Brain) ~ log(Gestation) + log(Litter) + log(Body), data=case0902)
summary(origMod)
```

5. **How do the coefficient estimates from the multiple imputation compare to the coefficient estimates on the original full data?**


## The coefficient estmates from the imputation are in the same ballpark as the original full data. Imputation via linear modeling in this case seems to be an over-estimate by roughly one standard error.
