---
title: "Module 8 Lab Examples"
output: pdf_document
---

```{r setup, include=FALSE}
library(boot)
library(ISLR)
library(leaps)
knitr::opts_chunk$set(echo = TRUE, exercise.eval = FALSE, warning = FALSE)
```

For this lab you'll follow along with the labs in *ISLR*.  Excerpts of the relevant sections are provided below. If you read through the same sections in the book, you'll notice the authors have a slightly different coding style (`=` for assignment instead of `<-`, `.` in variable names instead of `_`).  We have edited the assignment operator to match our usual style. You generally want to pick a style and stick with it, but of course there is no guarantee everyone else will use your style!

# Evaluating prediction accuracy

Follow along with sections 5.3.1, 5.3.2 and 5.3.3 in the section *5.3 Lab: Cross-Validation and the Bootstrap* in *ISLR*.

Take particular note of the functions:

* `poly()` - for an easy way to add polynomial terms to a regression model.
* `sample()` - as a way to randomly select row indexes of a data frame to extract data for a training set (and to exclude rows to get a test set).
* `cv.glm()` in the `boot` package to calculate leave-one-out and k-fold cross validation measures of error.


# The following two sections are excerpted and very slightly modified from *ISLR*. 

In this lab, we explore the resampling techniques covered in this chapter. Some of the commands in this lab may take a while to run on your computer.

## 5.3.1  The Validation Set Approach

We explore the use of the validation set approach in order to estimate the test error rates that result from fitting various linear models on the `Auto` data set. Before we begin, we use the `set.seed()` function in order to set a seed for R’s random number generator, so that the reader of this book will obtain precisely the same results as those shown below. It is generally a good idea to set a random seed when performing an analysis such as cross-validation that contains an element of randomness, so that the results obtained can be reproduced precisely at a later time. We begin by using the `sample()` function to split the set of observations into two halves, by selecting a random subset of 196 observations out of the original 392 observations. We refer to these observations as the training set.

```{r}
library(ISLR)
set.seed(1)
train <- sample(392, 196)
```

(Here we use a shortcut in the sample command; see `?sample` for details.)

We then use the `subset` option in `lm()` to fit a linear regression using only the observations corresponding to the training set.

```{r}
lm.fit <- lm(mpg ~ horsepower, data=Auto, subset=train)
```
We now use the `predict()`function to estimate the response for all 392 observations, and we use the `mean()` function to calculate the MSE of the 196 observations in the validation set. Note that the `-train` index below selects only the observations that are not in the training set.

```{r}
attach(Auto)
mean((mpg-predict(lm.fit,Auto))[-train]^2)
```
Therefore, the estimated test MSE for the linear regression fit is 26.14. We can use the `poly()` function to estimate the test error for the polynomial and cubic regressions.

```{r}
lm.fit2 <- lm(mpg ~ poly(horsepower, 2), data=Auto, subset=train)
mean((mpg-predict(lm.fit2, Auto))[-train]^2)

lm.fit3 <- lm(mpg ~ poly(horsepower, 3), data=Auto, subset=train)
mean((mpg-predict(lm.fit3, Auto))[-train]^2)
```

These error rates are 19.82 and 19.78, respectively. If we choose a different training set instead, then we will obtain somewhat different errors on the validation set.

```{r}
set.seed(2)
train <- sample(392,196)

lm.fit <- lm(mpg ~ horsepower, subset=train)
mean((mpg-predict(lm.fit, Auto))[-train]^2)

lm.fit2 <- lm(mpg ~ poly(horsepower, 2), data=Auto, subset=train)
mean((mpg-predict(lm.fit2, Auto))[-train]^2)

lm.fit3 <- lm(mpg ~ poly(horsepower, 3), data=Auto, subset=train)
mean((mpg-predict(lm.fit3, Auto))[-train]^2)
```


Using this split of the observations into a training set and a validation set, we find that the validation set error rates for the models with linear, quadratic, and cubic terms are 23.30, 18.90, and 19.26, respectively. These results are consistent with our previous findings: a model that predicts `mpg` using a quadratic function of `horsepower` performs better than a model that involves only a linear function of `horsepower`, and there is little evidence in favor of a model that uses a cubic function of horsepower.

## 5.3.2 Leave-One-Out Cross-Validation 

The LOOCV estimate can be automatically computed for any generalized linear model using the `glm()` and `cv.glm()` functions. In the lab for Chapter 4, we used the `glm()` function to perform logistic regression by passingin the `family="binomial"` argument. But if we use `glm()` to fit a model without passing in the `family` argument, then it performs linear regression, just like the `lm()` function. So for instance, 

```{r}
glm.fit <- glm(mpg ~ horsepower, data=Auto)
coef(glm.fit)
```
and
```{r}
lm.fit <- lm(mpg ~ horsepower, data=Auto)
coef(lm.fit)
```
yield identical linear regression models. In this lab, we will perform linear regression using the `glm()` function rather than the `lm()` function because the latter can be used together with `cv.glm()`. The `cv.glm()` function is part of the `boot` library.

```{r}
library(boot)
glm.fit <- glm(mpg ~ horsepower, data=Auto)
cv.err <- cv.glm(Auto, glm.fit)
cv.err$delta
```
The `cv.glm()` function produces a list with several components. The two numbers in the delta vector contain the cross-validation results. In this case the numbers are identical (up to two decimal places) and correspond to the LOOCV statistic given in (5.1). Below, we discuss a situation in which the two numbers differ. Our cross-validation estimate for the test error is approximately 24.23. We can repeat this procedure for increasingly complex polynomial fits. To automate the process, we use the `for()` function to initiate a *for loop* which iteratively fits polynomial regressions for polynomials of order `i=1` to `i= 5`, computes the associated cross-validation error, and stores it in the ith element of the vector `cv.error`. We begin by initializing the vector. This command will likely take a couple of minutes to run.

```{r} 
cv.error <- rep(0,5) 
for(i in 1:5){
  glm.fit <- glm(mpg ~ poly(horsepower,i), data=Auto)
  cv.error[i] <- cv.glm(Auto, glm.fit)$delta[1]
}
cv.error
```

As in Figure 5.4, we see a sharp drop in the estimated test MSE between the linear and quadratic fits, but then no clear improvement from using higher-order polynomials.

## 5.3.3 $k$-Fold Cross-Validation

The `cv.glm()` function can also be used to implement $k$-fold CV. Below we use $k = 10$, a common choice for k, on the `Auto` data set. We once again set a random seed and initialize a vector in which we will store the CV errors corresponding to the polynomial fits of orders one to ten.

```{r}
set.seed(17)
cv.error.10 <- rep(0,10)
for(i in 1:10){
  glm.fit <- glm(mpg ~ poly(horsepower, i), data=Auto)
  cv.error.10[i] <- cv.glm(Auto, glm.fit, K=10)$delta[1]
}
cv.error.10
```
Notice that the computation time is much shorter than that of LOOCV. (In principle, the computation time for LOOCV for a least squares linear model should be faster than for $k$-fold CV, due to the availability of the formula (5.2) for LOOCV; however, unfortunately the `cv.glm()` function does not make use of this formula.) We still see little evidence that using cubic or higher-order polynomial terms leads to lower test error than simply using a quadratic fit. We saw in Section 5.3.2 that the two numbers associated with `delta` are essentially the same when LOOCV is performed. When we instead perform $k$-fold CV, then the two numbers associated with `delta` differ slightly. The first is the standard $k$-fold CV estimate, as in (5.3). The second is a bias-corrected version. On this data set, the two estimates are very similar to each other.


# Variable selection

Follow along with sections 6.5.1, 6.5.2 and 6.5.3 in the section *6.5 Lab 1: Subset Selection Methods* in *ISLR*.

You'll learn about the `regsubsets()` function in the `leaps` package that facilitates fitting many regression models.

## 6.5.1  Best Subset Selection

Here we apply the best subset selection approach to the `Hitters` data. We wish to predict a baseball player’s `Salary` on the basis of various statistics associated with performance in the previous year. First of all, we note that the `Salary` variable is missing for some of the players. The `is.na()` function can be used to identify the missing observations. It returns a vector of the same length as the input vector, with a `TRUE` for any elements that are missing, and a `FALSE` for non-missing elements. The `sum()` function can then be used to count all of the missing elements.

```{r} 
library(ISLR)
names(Hitters)
dim(Hitters)
sum(is.na(Hitters$Salary))
```

Hence we see that `Salary` is missing for 59 players. The `na.omit()` function removes all of the rows that have missing values in any variable.

```{r} 
Hitters <- na.omit(Hitters)
dim(Hitters)
sum(is.na(Hitters))
```
The `regsubsets()` function (part of the `leaps` library) performs best subset selection by identifying the best model that contains a given number of predictors, where *best* is quantified using RSS. The syntax is the same as for `lm()`. The `summary()` command outputs the best set of variables foreach model size.

```{r}
library(leaps)
regfit.full <- regsubsets(Salary ~ ., Hitters)
summary(regfit.full)
```
An asterisk indicates that a given variable is included in the corresponding model. For instance, this output indicates that the best two-variable model contains only `Hits` and `CRBI`. By default, `regsubsets()` only reports results up to the best eight-variable model. But the `nvmax` option can be used in order to return as many variables as are desired. Here we fit up to a 19-variable model.
```{r} 
regfit.full <- regsubsets(Salary ~ ., data=Hitters, nvmax=19)
reg.summary <- summary(regfit.full)
```
The `summary()`function also returns $R^2$, RSS, adjusted $R^2$, $C_p$, and BIC. We can examine these to try to select the best overall model.
```{r}
names(reg.summary)
```
For instance, we see that the $R^2$ statistic increases from 32%, when only one variable is included in the model, to almost 55%, when all variables are included. As expected, the $R^2$ statistic increases monotonically as more variables are included.
```{r}
reg.summary$rsq
```
Plotting RSS, adjusted $R^2$, $C_p$, and BIC for all of the models at once will help us decide which model to select. Note the `type="l"` option tells R to connect the plotted points with lines.

The `points()` command works like the `plot()` command, except that it puts points on a plot that has already been created, instead of creating a new plot. The `which.max()` function can be used to identify the location ofthe maximum point of a vector. We will plot a red dot to indicate the model with the largest adjusted $R^2$ statistic.


```{r}
par(mfrow=c(2,2), mar=c(4,4,1,1))
plot(reg.summary$rss, xlab="Number of Variables", ylab="RSS", type="l")
plot(reg.summary$adjr2, xlab="Number of Variables", ylab="Adjusted RSq", type="l")
which.max(reg.summary$adjr2)
points(11, reg.summary$adjr2[11], col="red", cex=2, pch=20)
```

In a similar fashion we can plot the $C_p$ and BIC statistics, and indicate the models with the smallest statistic using `which.min()`.

```{r}
plot(reg.summary$cp, xlab="Number of Variables", ylab="Cp", type="l")
which.min(reg.summary$cp) 
points(10, reg.summary$cp[10], col="red", cex=2, pch=20)
which.min(reg.summary$bic)
plot(reg.summary$bic, xlab="Number of Variables", ylab="BIC", type="l")
points(6, reg.summary$bic[6], col="red", cex=2, pch=20)
```
The `regsubsets()` function has a built-in `plot()` command which can be used to display the selected variables for the best model with a given number of predictors, ranked according to the BIC, $C_p$, adjusted $R^2$, or AIC. To find out more about this function, type  `?plot.regsubsets`.

```{r}
plot(regfit.full, scale="r2")
plot(regfit.full, scale="adjr2")
plot(regfit.full, scale="Cp")
plot(regfit.full, scale="bic")
```

The top row of each plot contains a black square for each variable selected according to the optimal model associated with that statistic. For instance, we see that several models share a BIC close to -150. However, the model with the lowest BIC is the six-variable model that contains only `AtBat`, `Hits`, `Walks`, `CRBI`, `DivisionW`, and `PutOuts`. We can use the `coef()` function to see the coefficient estimates associated with this model.
```{r} 
coef(regfit.full, 6)
```

## 6.5.2  Forward and Backward Stepwise Selection

We can also use the `regsubsets()` function to perform forward stepwise or backward stepwise selection, using the argument `method="forward"` or `method="backward"`.

```{r}
regfit.fwd <- regsubsets(Salary ~ ., data=Hitters, nvmax=19, method="forward")
summary(regfit.fwd)
regfit.bwd <- regsubsets(Salary ~ ., data=Hitters, nvmax=19, method="backward")
summary(regfit.bwd)
```

For instance, we see that using forward stepwise selection, the best one-variable model contains only `CRBI`, and the best two-variable model additionally includes `Hits`. For this data, the best one-variable through six-variable models are each identical for best subset and forward selection. However, the best seven-variable models identified by forward stepwise selection, backward stepwise selection, and best subset selection are different.
```{r}
coef(regfit.full, 7)
coef(regfit.fwd, 7)
coef(regfit.bwd, 7)
```

## 6.5.3  Choosing Among Models Using the Validation Set Approach and Cross-Validation


We just saw that it is possible to choose among a set of models of different sizes using $C_p$, BIC, and adjusted $R^2$. We will now consider how to do this using the validation set and cross-validation approaches. In order for these approaches to yield accurate estimates of the test error, we must use only the training observations to perform all aspects of model-fitting--including variable selection. Therefore, the determination of which model of a given size is best must be made using only the training observations. This point is subtle but important. If the full data set is used to perform the best subset selection step, the validation set errors and cross-validation errors that we obtain will not be accurate estimates of the test error. In order to use the validation set approach, we begin by splitting the observations into a training set and a test set. We do this by creating a random vector, `train`, of elements equal to `TRUE` if the corresponding observation is in the training set, and `FALSE` otherwise. The vector `test` has a `TRUE` if the observation is in the test set, and a `FALSE` otherwise. Note the `!` in the command to create `test` causes `TRUE`s to be switched to `FALSE`s and vice versa. We also set a random seed so that the user will obtain the same training set/test set split.
```{r}
set.seed(1)
train <- sample(c(TRUE,FALSE), nrow(Hitters), replace=TRUE)
test <- (!train)
```

Now, we apply `regsubsets()` to the training set in order to perform best subset selection.
```{r}
regfit.best=regsubsets(Salary ~ ., data=Hitters[train,], nvmax=19)
```

Notice that we subset the `Hitters` data frame directly in the call in order to access only the training subset of the data, using the expression `Hitters[train,]`. We now compute the validation set error for the best model of each model size. We first make a model matrix from the `test` data.
```{r}
test.mat <- model.matrix(Salary ~ ., data=Hitters[test,])
```
The `model.matrix()` function is used in many regression packages for building an "X" matrix from data. Now we run a loop, and for each size `i`, we extract the coefficients from `regfit.best` for the best model of that size, multiply them into the appropriate columns of the test model matrix to form the predictions, and compute the test MSE.

```{r}
val.errors <- rep(NA,19) 
for(i in 1:19){
  coefi <- coef(regfit.best, id=i)
  pred <- test.mat[,names(coefi)] %*% coefi
  val.errors[i] <- mean((Hitters$Salary[test]-pred)^2)
}
```

We find that the best model is the one that contains ten variables.
```{r}
val.errors
which.min(val.errors)
coef(regfit.best,10)
```

This was a little tedious, partly because there is no `predict()` method for `regsubsets()`. Since we will be using this function again, we can capture our steps above and write our own predict method.
```{r}
predict.regsubsets <- function(object, newdata, id, ...){
  form <- as.formula(object$call[[2]])
  mat <- model.matrix(form, newdata)
  coefi <- coef(object,id=id)
  xvars <- names(coefi)
  mat[,xvars]%*%coefi
}
```

Our function pretty much mimics what we did above. The only complex part is how we extracted the formula used in the call to `regsubsets()`. We demonstrate how we use this function below, when we do cross-validation. Finally, we perform best subset selection on the full data set, and select the best ten-variable model. It is important that we make use of the full data set in order to obtain more accurate coefficient estimates. Note that we perform best subset selection on the full data set and select the best ten-variable model, rather than simply using the variables that were obtained from the training set, because the best ten-variable model on the full dataset may differ from the corresponding model on the training set.
```{r}
regfit.best <- regsubsets(Salary ~ ., data=Hitters, nvmax=19)
coef(regfit.best, 10)
```
In fact, we see that the best ten-variable model on the full data set has a different set of variables than the best ten-variable model on the training set. We now try to choose among the models of different sizes using cross-validation. This approach is somewhat involved, as we must perform best subset selection within each of the $k$ training sets. Despite this, we see that with its clever subsetting syntax, R makes this job quite easy. First, we create a vector that allocates each observation to one of $k=10$ folds, and we create a matrix in which we will store the results.
```{r}
k <- 10
set.seed(1)
folds <- sample(1:k, nrow(Hitters), replace=TRUE)
cv.errors <- matrix(NA, k, 19, dimnames=list(NULL, paste(1:19)))
```
Now we write a `for` loop that performs cross-validation. In the $j$th fold, the elements of folds that equal $j$ are in the test set, and the remainder are in the training set. We make our predictions for each model size (using our new `predict()` method), compute the test errors on the appropriate subset, and store them in the appropriate slot in the matrix `cv.errors`.
```{r}
for(j in 1:k){
  best.fit <- regsubsets(Salary ~ ., data=Hitters[folds!=j,], nvmax=19)
  for(i in 1:19){
    pred <- predict(best.fit,Hitters[folds==j,], id=i)
    cv.errors[j,i] <- mean((Hitters$Salary[folds==j]-pred)^2)
  }
}
```
This has given us a 10×19 matrix, of which the $(i, j)$th element corresponds to the test MSE for the $i$th cross-validation fold for the best $j$-variablemodel. We use the `apply()` function to average over the columns of this `apply()` matrix in order to obtain a vector for which the $j$th element is the cross-validation error for the $j$-variable model.
```{r}
mean.cv.errors <- apply(cv.errors ,2,mean)
mean.cv.errors
par(mfrow=c(1,1))
plot(mean.cv.errors, type='b')
```
We see that cross-validation selects an 11-variable model. We now perform best subset selection on the full data set in order to obtain the 11-variable model.
```{r}
reg.best <- regsubsets(Salary ~ ., data=Hitters, nvmax=19)
coef(reg.best, 11)
```

# Many models (optional)

The R code in the [Many Models](http://r4ds.had.co.nz/many-models.html) chapter in *R for Data Science* is quite advanced.  You should follow along with code but do not get too distracted by  all the details.  If you are feeling adventurous, after replicating the analysis in the reading, try to replicate the analysis in the lecture.  
