---
title: "ST517-HW1"
author: Ben Tankus
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Sleuth3)
library(ggplot2)

```


1. (3 points) (Adapted from Exercise 21, Section 6.7, The Statistical Sleuth, 2nd Ed.) The dataset
bearings.csv contains failure times (measured in millions of cycles) of engine bearings made from five
different compounds.

(a) (1 point) Read in the data. What type of data object is bearings? With ggplot2, create side-by-side boxplots of the failure times by compound.

```{r}
bearing <- read.csv(file = 'bearings.csv')
print(paste("data is stored in a", typeof(bearing)))
qplot(Compound, Time , data = bearing, geom = 'boxplot')

#length(bearing$Time[bearing$Compound == 'III'])


```

(b) (2 points) Determine the pairs of engine ball bearing compounds for which there is a significant
difference in mean failure times. Present your findings in a short statistical report ( 4 sentences).

$$H_0: \mu_{I} = \mu_{II},  \mu_{I} = \mu_{III}, \mu_{I} = \mu_{IV}, \mu_{I} = \mu_{V} ... for \ all \ pairs $$
$$H_A: \mu_{I} \neq \mu_{II},  \mu_{I} \neq \mu_{III}, \mu_{I} \neq \mu_{IV}, \mu_{I} \neq \mu_{V} ... for \ any \ pairs $$


```{r}
# Equal var, false. paired = false, independant = true, outliers = true, equal N = True
# METHOD - Use Tukey Kramer adjustment as we are doing all pairwise comparisons
fit <- aov(Time ~ Compound, data = bearing)
summary(fit)
TukeyHSD(fit)

```
Significant V - II & V - III

## There were two comparisons that reject the null hypothesis of equal means, V - II (p-val = 0.0008), and V - III (p-val = 0.031). Looking at the Box Plot, one would expect more pairs with statistical significance (V-IV, IV-II), but the high variance of V and the outlier of IV make comparisons less likely to be rejected. ANOVA procedures are not robust to outliers, and a high variance makes a sample much less likely to reject the null hypothesis. Due to the manufacturing quality nature of this testing and the strength of the outlier in IV, I reccomend removing the outlier in IV and retesting that sample to improve accuracy of the test.

2. (3 points) Using the data from the last weeks lab and homework, case0501 in the Sleuth3, answer the
question “Which diets differ in their mean lifetime?”

```{r}
dietData <- case0501
# METHOD - Use Tukey Kramer adjustment as we are doing all pairwise comparisons
#head(dietData)
fitDiets <- aov(Lifetime ~ Diet,data = dietData)

diets <- TukeyHSD(fitDiets)

diets
```

## Diets that Differ:
 N/R40-N/N85   12.4254386   8.885436  15.9654413 0.0000000  
 N/R50-N/N85   9.6059550   6.202170  13.0097399 0.0000000  
 NP-N/N85     -5.2891873  -9.017748  -1.5606269 0.0008380  
 R/R50-N/N85  10.1944862   6.593417  13.7955556 0.0000000  
 lopro-N/N85   6.9944862   3.393417  10.5955556 0.0000008  
 NP-N/R40    -17.7146259 -21.399845 -14.0294069 0.0000000  
 lopro-N/R40  -5.4309524  -8.987127  -1.8747778 0.0002306  
 NP-N/R50    -14.8951423 -18.449713 -11.3405719 0.0000000  
 R/R50-NP     15.4836735  11.739756  19.2275913 0.0000000  
 lopro-NP     12.2836735   8.539756  16.0275913 0.0000000  


3. (2 points) A soda company is developing a new soda. They are tying to determine how much sugar to
put in it to give it the best taste. In order to evaluate this, they have made samples with ten different
sugar levels. Each level of sugar is a assigned to a random sample of seven people, and each person rates
the soda on a scale from 0 to 10. The company would like to make inference on the difference between
mean ratings between each pair of sugar level.

(a) How many pairwise comparisons are there?

## There are 10c2 comparisons, or *45 combinations*

(b) Name 2 procedures that the company could use to control the familywise Type I error rate on the
differences of means? Explain.

## I reccomend using either the *TukeyHSD*, or *Bonferoni Adjustment*.


4. (2 points) A consumer research group names seven types of department stores. They take a random
sample of six department stores for each type and record their yearly sales. They wish to find significant
pairwise differences in mean yearly sales for seven types of stores.

(a) How many pairwise comparisons are there?

## There are 7c2 comparisons, or *21 combinations*

(b) The group wishes to control the familywise Type I error rate at 1 Percent using Bonferroni methods.
What should be the Type I error rate of each pairwise comparison?

## The type I error rate should be:
$$(1-\alpha/21)$$ , or $$ (1-0.01/21) $$ 
Which Equals *0.9995238*
